---
title: "Step1"
author: "ANV"
date: "May 31, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(data.table)
library(GEOquery)
library(stringdist)
source("Code/misc.R")

```

```{r}
# Specify a directory where GSE downloaded files are stored
GEOtemp <- "GEOtemp"

# Input file
GSELIST <- "test2.txt"
# Location/name for output file
REVIEW1 <- "sample_review.txt"
```

## Download GSEs
Feed in the ids of GSEs that passed Filter #1 (based on description). The following steps pull the metadata create a standardized annotation table for the samples.
```{r}
gses <- readLines(GSELIST)

# Download the GSE
getSamples <- function(gse) {
  eset <- getGEO(GEO = gse, destdir = GEOtemp)[[1]]
  pData(eset)
}

# if processing many, limit messages to avoid cluttering display
results <- suppressMessages(lapply(gses, function(x) try(getSamples(x)))) 
names(results) <- gses

```

Sometimes downloading a GEO will fail (maybe 1-5% of GSEs?) because there was a parsing error or some other problem. GEOquery verbose output but not always helpful, so better to check with:
```{r}

ok <- allOK(results)
table(ok)
cat("GSEs that failed:", names(results)[!ok])
results <- results[ok] # Go back to failures later, proceed with OK

```

## Create and populate annotation table
Infer some annotation information for each GSE using some simple rules. 
This is only the starting point as data is imperfectly extracted; manual annotation follows for the exported table. 
```{r}

# Experiment Type (1-Channel or 2-Channel; more specific assignments, i.e. whether there was dye swap, happens in human review)
inferXPType <- function(DT) {
  default <- rep("1C", nrow(DT))
  ch2 <- grep("source_name_ch2", names(DT))
  if(length(ch2)) return(rep("2C", nrow(DT)))
  return(default)
}

ixpType <- sapply(results, inferXPType)

# BSM -- Molecule -- note: perhaps have a list of all known tlr ligands to match against
inferBSM <- function(DT) {
  x1 <- rep("", nrow(DT))
  bsmcol <- grep("agent|stimulation", names(DT))
  if(length(bsmcol) == 1) {
    info <- DT[bsmcol]
    return(info)
  } 
  return(x1)
}

iBSM <- sapply(results, inferBSM)

# BSMDCD -- Time
inferDCD <- function(DT) {
  x1 <- rep("", nrow(DT))
  # First find a column that specifically contains treatment info
  tcol <- grep("^treatment:ch1|time", names(DT))
  if(length(tcol)) {
    info <- do.call(paste, DT[tcol])
  } else { # Use information in sample title
    info <- DT$title
  }
  m <- regexpr("[0-9]+ ?(d|h|D|H)|(D|d)(ay) ?[0-9]+", info)
  x1[m != -1] <- regmatches(info, m)
   # Sometimes time is stated as "overnight", which usually means "12 h"
  x1[grepl("overnight", info)] <- "12h"
  x1 <- tolower(gsub(" ", "", x1))
  return(x1)
}

iDCD <- sapply(results, inferDCD) 

# BSMDCD2 -- Dose/concentration
inferDCD2 <- function(DT) {
  x1 <- rep("", nrow(DT))
  dosecol <- grep("dose|concentration|amount|treatment", names(DT))
  dosecol <- c(dosecol, which(sapply(DT, function(x) grepl("treatment_protocol", x[1])) == T))
  if(length(dosecol)) {
    info <- do.call(paste, DT[dosecol])
    m <- gregexpr("[0-9.]+ ?(u|n|m|[?])(g|M)?(\\/[A-Za-z]{2})?", info)
    dose <- sapply(regmatches(info, m), paste, collapse = ";")
    dose <- gsub("?", "u", dose, fixed = T)
    x1 <- dose
  }
  return(x1)
}

iDCD2 <- sapply(results, inferDCD2) 

# BiosampName -- to do: a more sophisticated implementation could also match to biosample ID
biosamps <- readLines("biosamps.txt") ## This is a list derived from "allbiosamples"

inferBiosamp <- function(DT) {
  # First find the obvious columns containing cell line/type
  cellcol <- grep("^cell.*ch1", colnames(DT))
  if(length(cellcol)) {
    cell <- do.call(paste, DT[cellcol])
    return(cell)
  } else {
    tissuecol <- grep("^tissue.*ch1", colnames(DT))
    if(length(tissuecol)) {
      tissue <- do.call(paste, DT[tissuecol])
      return(tissue)
    } else { # Use the source_name_ch1,treatment_protocol_ch1 columns otherwise
      bs <- lapply(biosamps, function(b) grepl(b, paste(DT$source_name_ch1, DT$treatment_protocol_ch1)))
      names(bs) <- biosamps
      bs <- bs[sapply(bs, sum) >= 1 ]
      if(!length(bs)) return(rep("?", nrow(DT))) 
      bs <- bs[order(sapply(bs, sum), decreasing = T)]
      nterm <- ifelse(length(bs) > 1, 2, 1)
      x1 <- lapply(1:nterm, function(i) ifelse(bs[[i]], names(bs)[[i]], ""))
      x1 <- do.call("paste", x1)
      x1 <- trimws(x1)
      return(x1)
    }
  }
}

iBiosamp <- sapply(results, inferBiosamp)

# Control?
inferCTRL <- function(DT) {
  s <- DT$title
  ctrl <- as.numeric(grepl("wt|wildtype|control|ctrl|healthy|media|medium|veh|unstim|untreated|mock", s, ignore.case = T))
}

iCTRL <- sapply(results, inferCTRL)

```


Subset table for export with selected columns. Join data in one table and fill in additional columns.
```{r}

export <- rbindlist(results, use.names = T, fill = T)
keepcols <- grep("title|geo_accession|source_name|description|platform|characteristics|treatment", names(export))
export <- export[, keepcols, with = F]
export[, GSE := unlist(mapply(rep, names(results), sapply(results, nrow)))]
newcols1 <- c("RefDye", "Comment", "Group", "Batch", "Node", "NodeFunction", "BSM") 
export[, (newcols1) := ""]
export[, c("Ignore") := 0]
export[, c("XP") := 1]

# Inferred variables
export[, xpType := unlist(ixpType)]
export[, isCTRL := unlist(iCTRL)]
export[, BioSampName := unlist(iBiosamp)]
export[, BSMDCD := unlist(iDCD)]
export[, BSMDCD2 := unlist(iDCD2)]

myorder <- c("GSE", "xpType", "RefDye", "Comment", "Group", "Ignore", "XP", "isCTRL", "Batch", "Node", "NodeFunction", "BSM", "BSMDCD", "BSMDCD2", "BioSampName",
              "title", "geo_accession")
neworder <- c(myorder, names(export)[!names(export) %in% myorder])
setcolorder(export, neworder)

head(export)
write.table(export, REVIEW1, sep = "\t", row.names = F, quote = F)

```

